### 2025-05-05 Chapters/Chapter4.tex
Bætti Stórhöfða við, uppfærði síðasta kaflann. Tók út * fyrir undirkafla og eitthvað. Notaði svo GPT til þess að benda mér á stafsetningarvillur og uppfærði. Held að þetta sé nokkuð langt komið.

As the average wind speed increases, then the variability in gust as a percentage decreases\cite{mean_gust_HA_HO}. Therefore lower MAPE woult be expected if the training data is restricted to higher wind speeds. This fact explains why several models were created for different lower bounds.

\section{Modeling for CARRA wind speed above 10 m/s}
As noted in Chapter 1, it is more important to predict accurate gust factor when it is windy than when the weather is calm. Therefore, at first, the data is restricted to ws$_{15} \geq 10$ m/s. The first constructed model incorporates the features temperature, pressure and wind direction at 15 m height. This gives a moderately reduced MAPE of 16.7\% compared with 17.2\% for the baseline ws$_{15}$ model. When the next pair of features, station altitude and transformed wind direction, is included, the MAPE is reduced considerably, down to 13.8\%. The atmospheric stability indicators, Ri and $N^2$, add little to the model performance, contrary to what might have been expected. When the atmosphere is unstable, wind gusts from higher up might be expected. However, another considerable perfomance increase is observed when digital elevation model (DEM) information is included, down to a MAPE of 12.0\%. This represents over 30\% reduction in the MAPE from the ws$_{15}$ baseline model.

\begin{table}[h]
    \caption[Model results for different sets of parameters.]{Showing the results of different models. Models defined by different sets of parameters, with increasing complexity. This table shows the feature importance as the loss decreases as more variables are added. Adding all height levels decreases the loss but very little.}
    \label{table:setsOfParams}
    \centering
    \begin{tabular}{lc}
        \toprule
        \textbf{Model variables} & \textbf{MAPE}\\
        \midrule
        Baseline constant & 23.9\%\\
        $ws_{15}$ & 17.2\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}]$ & 16.7\% \\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, \text{altitude}, twd]$ & 13.8\% \\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, \text{altitude}, twd, N^2, Ri]$ & 13.7\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, \text{altitude}, twd, N^2, Ri]$ + DEM & 12.0\%\\
        $[ws_{15,250,500}, t_{15,250,500}, p_{15,250,500}, wd_{15,250,500}, \text{altitude}, twd, N^2, Ri]$ & 12.0\% \\
        \bottomrule
    \end{tabular}
\end{table}

The results show this. Several different AWSL were used for the baseline model, as can be seen in table \ref{table:results}. This sets a goal. A model that does not significantly improve on this baseline suggests either failure to capture essential patterns in the data or that the data itself may lack the necessary information for substantial improvements upon the baseline. Using the previously described neural network architecture setups for each AWSL, with and without landscape elevation information, MAPE was determined. The results can be seen in Table \ref{table:results}.

\begin{table}[h]
    \caption[Model results for different AWSL]{MAPE for each average wind speed limit with and without landscape elevation in a 30° sector around the point of interest into the direction of the reanalysis wind. The influence of adding elevation data seems to reduce the error. The percentage error is higher for lower wind speeds and thus observing the error for different lower bound of wind speed will produce different results. This lower bound is determined using the reanalysis wind speed at 15 meters ($ws_{15}$).}
    \label{table:results}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \textbf{AWSL} & \textbf{MAPE} & &\\ 
        $[m/s]$ & \textit{Baseline} &  \textit{Without DEM} & \textit{With DEM} \\
        \midrule
        $\geq 0$ & 39.2\% & 19.2\% & 18.9\% \\
        $\geq 5$ & 28.1\% & 15.3\% & 14.9\%\\
        $\geq 10$ & 23.9\% & 13.3\% & 12.5\%\\
        $\geq 15$ & 23.2\% & 14.4\% & 11.9\%\\
        $\geq 20$ & 24.7\% & 15.7\% & 13.3\%\\
        $\geq 25$ & 27.7\% & 19.4\% & 17.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result by station]{The MAPE results for selected stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \textbf{Station name} & \textbf{\# points} & \multicolumn{2}{c}{\textbf{MAPE}}\\
        & & General training & Site training\\
        \midrule
        Akrafjall & 42,791 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,014 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,121 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,176 & 7.7\% & 7.7\%\\
        Sandbúðir & 18,718 & 6.8\% & 6.4\%\\
        Stórholt & 35,126 & 7.1\% & 29.2\% \\
        Þúfuver & 19,538 & 6.4\% & 6.8\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result looking at closed wind speed intervals]{The MAPE results for different AWSL intervals. Here instead of training for all data above a certain threshold put of the wind, training is done only on data between two wind speeds. The percentage variance in gust factor as a function of wind speed increases with decreasing wind speed. Measured wind speed is used for the cutoff and thus have data leakage. This results should thus be somewhat comparable to the last column in Table (\ref{table:results})}
    \label{table:closed_intervals}
    \centering
    \begin{tabular}{lcc}
        \toprule
        \textbf{Interval} &  \multicolumn{2}{c}{\textbf{MAPE}}\\
        $[m/s]$ & Without Elevation & With Elevation\\
        \midrule
        $[5, 10[$ & 17.1\% & 16.4\%\\
        $[10, 15[$ & 14.5\% & 13.0\%\\
        $[15, 20[$ & 15.0\% & 12.0\%\\
        $[20, 25[$ & 15.6 \% & 13.1\%\\
        $[25, 30[$ & 18.4\% & 19.0\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result by stations of interest]{The MAPE results of different stations for several stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the AWSL is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:more_specific_sites}
    \centering
    \begin{tabular}{lcc}
        \toprule
        \textbf{Station name} & \multicolumn{2}{c}{\textbf{MAPE}}\\
         & Baseline & Model\\
        \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

In Figure (\ref{fig:ShapleySummary}) the contribution of each feature, excluding the elevation points, can be seen for the model in general (a significant subset of data is used). Looking at Figure (\ref{fig:ShapleySummary}), there is an outlier. Exactly calculating the Shapley values is time intensive, in the subset shown there is an outlier that skews the figure and makes it so that viewing the importance distribution excluding the outlier is difficult. For this reason, another shapley summary vas created that looked at different, and larger distribution. This can be seen in Figure (\ref{fig:ShapleySummary2}). The Figures (\ref{fig:ShapleySummary}) and (\ref{fig:ShapleySummary2}) show the summary for a model trained only on the features shown and not on landscape elevation of the surrounding area. This is done as the number of points there is too high to display in one figure (70 total points). Looking specifically at Figure (\ref{fig:ShapleySummary2}), the station elevation is most influential and the Richardson number's influence is very low. Most of the feature values bunch up in the middle, while the station elevation is elongated compared to other features. Station elevation seems to have two bunches on either side of 0. Overall, the values seem to be skewed to the right of 0. This would be expected as the predicted values are expected to be in the range of 1.2-2, or at least always above 1 by definition. Finally, for the Shapley values, looking that all the data there are again outliers that skew the data so that spotting general distribution is difficult. This can be seen in Figure (\ref{fig:ShapleySummary3}).

@@

Further Shapley figures can be seen in Appendix~\ref{appendix:A}. In each of these plots, even without the feature labels, the station elevation is easily noticed as the value is constant for a station. This is not noteworthy. What is noteworthy is the range of impact from this single value. For simpler models, this would not happen. It is important to note that SHAP assumes feature independece\cite{Salih_2024}. This might explain why the impact of the Richardson number is so low. Both the squared Brunt–Väisälä frequency and the Richardson number are derived features from reanalysis data. They carry with them some extra information over the other features in the dataset. This is because both are variables over elevation ranges. That is, as seen in Equations (\ref{eqn:Ri}, \ref{eqn:N}), both are dependent on values at lower and upper elevations and try to describe the stability of that range. Shapley tries to assign contribution values for each feature for each observation. SHAP assumes that the features are independent, but this is not the case. It is clearly not the case for the derived variables, but how the contribution should be distributed between the features is not clear. Seemingly the SHAP python package is giving all the impact to the Brunt–Väisälä squared frequency and none to Richardson number. If the Brunt–Väisälä would be excluded from the data, the impact of the Richardson number would likely increase. Another point to note is that the features are ordered by their impact. This means that the station elevation is the most impactful for each plot, but the ordering of other variables changes. Looking at Figure (\ref{fig:ShapleySummary3}), which shows the Shapley summary plot for all data, the wind speed is the second most important feature. This is reversed in Figure (\ref{fig:ShapleySummaryAkrafjall}). This is not unexpected as Akrafjall station was specifically selected as the MAE for reanalysis wind speed was very high as can be seen in Table (\ref{fig:ShapleySummaryAkrafjall}), where you will also find the stations whose summary plot is shown in Figures (\ref{fig:ShapleySummaryAlmannaskarð}, \ref{fig:ShapleySummaryAsgarðsfjall}) and these also fall into the category of very high MAE for wind speed. What is interesting is that the reanalysis wind speed is also of low impact at stations like Háahlíð and Keflavíkurflugvöllur, as shown in Figures (\ref{fig:ShapleySummaryHaahlid}, \ref{fig:ShapleySummaryKeflavikurflugvollur}). These stations had the lowest of MAE for difference between measured wind speed and reanalysis wind speed. As the summary plot over all stations (Figure (\ref{fig:ShapleySummary3})) shows that reanalysis wind speed is impactful, might lead to the conclusion that the reanalysis wind speed is not a good predictor at these locations or something else is skewing the data. A simpler way to look at feature importance is to create models that are trained on and use to predict different sets of parameters. The results of such a comparison can be seen in Table \ref{table:setsOfParams}.

>>>>

As the average wind speed increases, then the variability in gust as a percentage decreases\cite{mean_gust_HA_HO}. Therefore lower MAPE would be expected if the training data is restricted to higher wind speeds. This fact explains why several models were created for different lower bounds.

\section{Modeling for CARRA wind speed above 10 m/s}
As noted in Chapter \ref{Chapter1}, it is more important to predict accurate gust factor when it is windy than when the weather is calm. Therefore, at first, the data is restricted to ws$_{15} \geq 10$ m/s. The first constructed model incorporates the features temperature, pressure and wind direction at 15 m height. This gives a moderately reduced MAPE of 16.7\% compared with 17.2\% for the baseline ws$_{15}$ model. When the next pair of features, station altitude and transformed wind direction, is included, the MAPE is reduced considerably, down to 13.8\%. The atmospheric stability indicators, Ri and $N^2$, add little to the model performance, contrary to what might have been expected. When the atmosphere is unstable, wind gusts from higher up might be expected. However, another considerable performance increase is observed when digital elevation model (DEM) information is included, down to a MAPE of 12.0\%. This represents over 30\% reduction in the MAPE from the ws$_{15}$ baseline model.

\begin{table}[h]
    \caption[Model results for different sets of parameters.]{Showing the results of different models. Models defined by different sets of parameters, with increasing complexity. This table shows the feature importance as the loss decreases as more variables are added. Adding all height levels decreases the loss but very little.}
    \label{table:setsOfParams}
    \centering
    \begin{tabular}{lc}
        \toprule
        Model variables & MAPE\\
        \midrule
        Baseline constant & 23.9\%\\
        $ws_{15}$ & 17.2\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}]$ & 16.7\% \\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, \text{altitude}, twd]$ & 13.8\% \\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, \text{altitude}, twd, N^2, Ri]$ & 13.7\%\\
        $[ws_{15}, t_{15}, p_{15}, wd_{15}, \text{altitude}, twd, N^2, Ri]$ + DEM & 12.0\%\\
        $[ws_{15,250,500}, t_{15,250,500}, p_{15,250,500}, wd_{15,250,500}, \text{altitude}, twd, N^2, Ri]$ & 12.0\% \\
        \bottomrule
    \end{tabular}
\end{table}

The results show this. Several different wind speed intervals were used for the baseline model, as can be seen in table \ref{table:results}. This sets a goal. A model that does not significantly improve on this baseline suggests either failure to capture essential patterns in the data or that the data itself may lack the necessary information for substantial improvements upon the baseline. Using the previously described neural network architecture setups for each wind speed interval, with and without landscape elevation information, MAPE was determined. The results can be seen in Table \ref{table:results}.

\begin{table}[h]
    \caption[Model results for different wind speed limits]{MAPE for each average wind speed limit with and without landscape elevation in a 30° sector around the point of interest into the direction of the reanalysis wind. The influence of adding elevation data seems to reduce the error. The percentage error is higher for lower wind speeds and thus observing the error for different lower bound of wind speed will produce different results. This lower bound is determined using the reanalysis wind speed at 15 meters ($ws_{15}$).}
    \label{table:results}
    \centering
    \begin{tabular}{lccc}
        \toprule
        \shortstack{Wind Speed\\limit} & MAPE & &\\
        $[m/s]$ & \shortstack{Baseline\\constant} &  Without DEM & With DEM \\
        \midrule
        $\geq 0$ & 39.2\% & 19.2\% & 18.9\% \\
        $\geq 5$ & 28.1\% & 15.3\% & 14.9\%\\
        $\geq 10$ & 23.9\% & 13.3\% & 12.5\%\\
        $\geq 15$ & 23.2\% & 14.4\% & 11.9\%\\
        $\geq 20$ & 24.7\% & 15.7\% & 13.3\%\\
        $\geq 25$ & 27.7\% & 19.4\% & 17.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result by station]{The MAPE results for selected stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the wind speed limit is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:specific_sites}
    \centering
    \begin{tabular}{lccc}
        \toprule
        Station name & n & \multicolumn{2}{c}{MAPE}\\
        & & General training & Site training\\
        \midrule
        Akrafjall & 42,791 & 18.6\% & 93.7\%\\
        Almannaskarð & 4,014 & 12.2\% & 86.7\%\\
        Ásgarðsfjall & 15,121 & 9.1\% & 9.4\%\\
        Jökulheimar & 17,176 & 7.7\% & 7.7\%\\
        Sandbúðir & 18,718 & 6.8\% & 6.4\%\\
        Stórholt & 35,126 & 7.1\% & 29.2\% \\
        Þúfuver & 19,538 & 6.4\% & 6.8\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result looking at closed wind speed intervals]{The MAPE results for different wind speed limit intervals. Here instead of training for all data above a certain threshold of wind speed, training is done only on data between two wind speeds. The percentage variance in gust factor as a function of wind speed increases with decreasing wind speed. Measured wind speed is used for the cutoff and thus have data leakage. These results should thus be somewhat comparable to the last column in Table (\ref{table:results})}
    \label{table:closed_intervals}
    \centering
    \begin{tabular}{lcc}
        \toprule
        Interval &  \multicolumn{2}{c}{MAPE}\\
        $[m/s]$ & Without Elevation & With Elevation\\
        \midrule
        $[5, 10[$ & 17.1\% & 16.4\%\\
        $[10, 15[$ & 14.5\% & 13.0\%\\
        $[15, 20[$ & 15.0\% & 12.0\%\\
        $[20, 25[$ & 15.6 \% & 13.1\%\\
        $[25, 30[$ & 18.4\% & 19.0\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[h]
    \caption[Model result by stations of interest]{The MAPE results of different stations for several stations of interest, both when training for the specific site and when the stations are a part of the general data. For every station the wind speed limit is set at 10 m/s. In training for a single station at a time, some site specific information can be gauged. This does not mean that the a better result can be reached for that site. Factors such as the number of datapoints at given location can significantly impact the result. This table uses the measured wind speed to determine the cut off for data points. This leads to some data leakage and an increased performance compared to using the reanalysis CARRA speed for cut off.}
    \label{table:more_specific_sites}
    \centering
    \begin{tabular}{lcc}
        \toprule
        Station name & \multicolumn{2}{c}{MAPE}\\
         & Baseline & Model\\
        \midrule
        Fáskrúðsfjörður & 28.2\% & 21.8\%\\
        Ingólfsfjall & 30.0\% & 19.6\%\\
        Kjalarnes & 20.7\% & 13.5\% \\
        Sandskeið & 13.0\% & 10.2\%\\
        Seyðisfjörður & 32.1\% & 23.0\%\\
        Þjórsárdalur & 12.2\% & 11.4\%\\
        Þrengsli & 13.6\% & 11.3\%\\
        \bottomrule
    \end{tabular}
\end{table}

@@

In Figure (\ref{fig:ShapleySummary}) the contribution of each feature, excluding the elevation points, can be seen for the model in general (a significant subset of data is used). Looking at Figure (\ref{fig:ShapleySummary}), there is an outlier. Exactly calculating the Shapley values is time intensive, in the subset shown there is an outlier that skews the figure and makes it so that viewing the importance distribution excluding the outlier is difficult. For this reason, another shapley summary was created that looked at different, and larger distribution. This can be seen in Figure (\ref{fig:ShapleySummary2}). The Figures (\ref{fig:ShapleySummary}) and (\ref{fig:ShapleySummary2}) show the summary for a model trained only on the features shown and not on landscape elevation of the surrounding area. This is done as the number of points there is too high to display in one figure (70 total points). Looking specifically at Figure (\ref{fig:ShapleySummary2}), the station elevation is most influential and the Richardson number's influence is very low. Most of the feature values bunch up in the middle, while the station elevation is elongated compared to other features. Station elevation seems to have two bunches on either side of 0. Overall, the values seem to be skewed to the right of 0. This would be expected as the predicted values are expected to be in the range of 1.2-2, or at least always above 1 by definition. Finally, for the Shapley values, looking that all the data there are again outliers that skew the data so that spotting general distribution is difficult. This can be seen in Figure (\ref{fig:ShapleySummary3}).

@@

Further Shapley figures can be seen in Appendix~\ref{appendix:A}. In each of these plots, even without the feature labels, the station elevation is easily noticed as the value is constant for a station. This is not noteworthy. What is noteworthy is the range of impact from this single value. For simpler models, this would not happen. It is important to note that SHAP assumes feature independence\cite{Salih_2024}. This might explain why the impact of the Richardson number is so low. Both the squared Brunt–Väisälä frequency and the Richardson number are derived features from reanalysis data. They carry with them some extra information over the other features in the dataset. This is because both are variables over elevation ranges. That is, as seen in Equations (\ref{eqn:Ri}, \ref{eqn:N}), both are dependent on values at lower and upper elevations and try to describe the stability of that range. Shapley tries to assign contribution values for each feature for each observation. SHAP assumes that the features are independent, but this is not the case. It is clearly not the case for the derived variables, but how the contribution should be distributed between the features is not clear. Seemingly the SHAP python package is giving all the impact to the Brunt–Väisälä squared frequency and none to Richardson number. If the Brunt–Väisälä would be excluded from the data, the impact of the Richardson number would likely increase. Another point to note is that the features are ordered by their impact. This means that the station elevation is the most impactful for each plot, but the ordering of other variables changes. Looking at Figure (\ref{fig:ShapleySummary3}), which shows the Shapley summary plot for all data, the wind speed is the second most important feature. This is reversed in Figure (\ref{fig:ShapleySummaryAkrafjall}). This is not unexpected as Akrafjall station was specifically selected as the MAE for reanalysis wind speed was very high as can be seen in Table (\ref{fig:ShapleySummaryAkrafjall}), where you will also find the stations whose summary plot is shown in Figures (\ref{fig:ShapleySummaryAlmannaskarð}, \ref{fig:ShapleySummaryAsgarðsfjall}) and these also fall into the category of very high MAE for wind speed. What is interesting is that the reanalysis wind speed is also of low impact at stations like Háahlíð and Keflavíkurflugvöllur, as shown in Figures (\ref{fig:ShapleySummaryHaahlid}, \ref{fig:ShapleySummaryKeflavikurflugvollur}). These stations had the lowest of MAE for difference between measured wind speed and reanalysis wind speed. As the summary plot over all stations (Figure (\ref{fig:ShapleySummary3})) shows that reanalysis wind speed is impactful, might lead to the conclusion that the reanalysis wind speed is not a good predictor at these locations or something else is skewing the data. A simpler way to look at feature importance is to create models that are trained on and use to predict different sets of parameters. The results of such a comparison can be seen in Table \ref{table:setsOfParams}.

### 2025-05-05 Chapters/Chapter5.tex
Update Chapter5.tex
Breytti nafni á fimmta kafla.
Bætti Stórhöfða við, uppfærði síðasta kaflann. Tók út * fyrir undirkafla og eitthvað. Notaði svo GPT til þess að benda mér á stafsetningarvillur og uppfærði. Held að þetta sé nokkuð langt komið.

\chapter{Discussion}
\label{Chapter5}
The goal of this tesis was to research whether it was possible to use reanalysis data to increase the predictability of wind gusts and see what influence including the landscape elevation would have. A quick statistical analysis will give a p-value of 0, that is these results are statistically significant. The improvement shown in Table (\ref{table:results}) is not chance. There is some pattern to be learned in the interval that leads to a higher predictability of wind gusts over baseline. There is also a large variability in the predictability based on weather stations. A large part of that can be attributed to the differing amount of data points for each station. It would be interesting to know how much of the predictability difference can be explained away with more data for these stations that have few data points or if there is some inherent difference. If there is inherent less predictability at those stations, what could be contributing factors in that?

>>>>

\chapter{Discussion and conclusions}
\label{Chapter5}
As shown in Table \ref{table:results}, the neural-network model achieves a significant reduction in mean absolute percentage error (MAPE) compared to a constant baseline. Table \ref{table:setsOfParams} further demonstrates a meaningful improvement over a simple regression model. A statistical test yields $p<0.01$, confirming that these improvements are unlikely to be due to chance. The influence of individual features on model performance is illustrated by Shapley plots (Figure \ref{fig:ShapleySummaryKeflavikurflugvollur} and Appendix \ref{appendix:A}). Although one might expect stability indicators—such as the Richardson number and the Brunt--Väisälä frequency—to be among the strongest predictors of gust variability, they contributed less than other variables. For reanalysis wind speeds above 10~m/s, the network reduces error by about 30\% relative to the regression baseline. The largest gains appear in the 10--25~m/s range (Table \ref{table:closed_intervals}). At lower speeds, percentage errors inflate even when absolute gust--wind differences remain small; at higher speeds, sample sizes become sparse.

Predictive skill varies notably from station to station. This variability can be partly attributed to discrepancies between reanalysis and observations (Table \ref{table:station_mae_distribution}), differences in local terrain, and the number of measurements per site. Training separate models for individual stations did not improve performance (Tables \ref{table:specific_sites} and \ref{table:more_specific_sites}); in fact, some well-sampled stations performed worse when modeled in isolation. Future work should investigate whether these anomalies arise from systematic measurement errors, inherent atmospheric variability at those sites, or other factors.

This study did not include station coordinates (XY) as direct inputs. Adding location information could improve site-specific accuracy without losing the benefits of training on the full dataset. Likewise, instead of applying a fixed linear interpolation to CARRA and elevation grids, the network could receive raw neighboring grid values and their relative positions, allowing it to learn an optimal interpolation scheme.

Finally, treating the data as a time series may offer further gains. Here, only instantaneous inputs at forecast time were used, partly because CARRA fields are available every three hours and may not capture rapid stability changes. Access to higher-frequency data or the inclusion of temporal history could enable models to exploit evolving atmospheric conditions and further enhance gust-prediction skill.
